---
title: "FEL results"
author: "sadie"
date: "5/11/2020"
output: html_document
---
```{r setup, include=FALSE}
library(jsonlite)

library(stringr)

library(dplyr)
library(purrr)
library(plyr)
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization
library(dendextend) # for comparing two dendrograms
```

# check one alignment
## zeiformes nd1 FEL results
```{r}
filepath <- read_json("~/bin/mtDNA_redo/data/FEL/zeiformes-nd1-align-dna.fas.FEL.json") #read in json

heads <- filepath$MLE$headers  %>% unlist() %>% .[c(TRUE,FALSE)] #get headers and ignore header descriptions
```

```{r}
#get MLE contents and make them a data frame
temp <- filepath$MLE$content$`0` %>% unlist %>% matrix(ncol = 6, byrow = TRUE) %>% as.data.frame() 
#make the headers the variable names
 names(temp)  <- heads
```

#kernel density and plots?
##Kernel density and plot for alpha for single FEL result
```{r}
d_alpha <- density(log(temp$alpha), kernel = "gaussian", from = 0) #Kernel density for alpha rate estimation 
#start from 0 and log transform to control outliers

d_alpha %>% plot() #plot the density 
```
what is the variability of the distribution?
```{r}
d_a_var <- d_alpha$x %>% var()

# t.a <- temp$alpha
# t.a[which(temp$alpha<=0)]  <- 0 
# t.a[which(temp$alpha== -Inf)]  <- 0 
#   log(t.a)%>% var()
```

##Kernel density and plot for beta for single FEL result
```{r}
d_beta <- density(log(temp$beta), kernel = "gaussian", from = 0)  #Kernel density for the beta rate estimation

d_beta %>% plot() #plot the density
```


#check a second alignment:
## acipenseriformes atp8 FEL results
```{r}
filepath <- read_json("~/bin/mtDNA_redo/data/FEL/acipenseriformes-atp8-align-dna.fas.FEL.json") #read in json

heads <- filepath$MLE$headers  %>% unlist() %>% .[c(TRUE,FALSE)] #get headers and ignore header descriptions
```

```{r}
#get MLE contents and make them a data frame
temp_2 <- filepath$MLE$content$`0` %>% unlist %>% matrix(ncol = 6, byrow = TRUE) %>% as.data.frame() 
#make the headers the variable names
 names(temp_2)  <- heads
```

#kernel density and plots?
##Kernel density and plot for alpha for single FEL result
```{r}
d_alpha_2 <- density(log(temp_2$alpha), kernel = "gaussian", from = 0)

d_alpha_2 %>% plot()
```

what is the variability of the distribution?
```{r}
d_a_2_var <- d_alpha_2$x %>% var()
```

##Kernel density and plot for beta for single FEL result
```{r}
d_beta_2 <- density(log(temp_2$beta), from = 0)
d_beta_2 %>% plot()
```





```{r}
dir <- "~/bin/mtDNA_redo/data/FEL"
FEL_jsons <- list.files(path = dir,
                      pattern = '*FEL.json', recursive = TRUE, full.names = TRUE)
```

```{r}
d.temp <- NULL
densities <- function(filename){
  
  results <- read_json(filename) #read in json

  heads <- results$MLE$headers  %>% unlist() %>% .[c(TRUE,FALSE)] #get headers and ignore header descriptions
  
  #get MLE contents and make them a data frame
  temp <- results$MLE$content$`0` %>% unlist %>% matrix(ncol = 6, byrow = TRUE) %>% as.data.frame() 
#make the headers the variable names
  names(temp)  <- heads
  
#Kernel density for alpha rate estimation 
#start from 0 and log transform to control outliers
   d_alpha <- density(log(temp$alpha), kernel = "gaussian", from = 0) 
  # d_beta <- density(log(temp$beta), kernel = "gaussian", from = 0) 
   
  #dist.temp <- dist.JSD(temp %>% select(alpha, beta))
  d.temp <- cbind(filename,  temp, stringsAsFactors = FALSE)
  
  return(d.temp)
}
```

```{r}
densities(FEL_jsons[1]) -> x

do.call(rbind, lapply(FEL_jsons,densities))-> x

cleannames <- x$filename %>% str_extract_all(., "\\w+(?=-)", simplify = T)

x <- x %>% mutate(order = toupper(cleannames[,1]), gene = toupper(cleannames[,2]))
```

```{r}
x %>% filter(order == "ACIPENSERIFORMES") %>% ggplot() +geom_density(aes(x = log(alpha), color = gene)) + xlim(0,9)
```
```{r}
x %>% filter(order == "ACIPENSERIFORMES") %>% ggplot() +geom_density(aes(x = log(beta), color = gene)) + xlim(0,9)
```

```{r}
x %>% filter(gene == "ATP8") %>% ggplot() +geom_density(aes(x = log(alpha), color = order)) + xlim(0,9) +guides(color = FALSE) + ylim(0,3)
```


```{r}
x %>% ggplot() +geom_density(aes(x = log(alpha), color = gene)) + facet_wrap(vars(order), scales = "free") + guides(color = FALSE)
```
```{r}
x %>% ggplot() +geom_density(aes(x = log(alpha), color = order)) + facet_wrap(vars(gene), scales = "free") + guides(color = FALSE)
```

```{r}
x %>% write.csv("mtDNA_FEL_results.csv")
```

## zeiformes atp6 FEL results
```{r}
filepath <- read_json("~/bin/mtDNA_redo/data/FEL/zeiformes-atp6-align-dna.fas.FEL.json") #read in json

heads <- filepath$MLE$headers  %>% unlist() %>% .[c(TRUE,FALSE)] #get headers and ignore header descriptions
```

```{r}
#get MLE contents and make them a data frame
temp <- filepath$MLE$content$`0` %>% unlist %>% matrix(ncol = 6, byrow = TRUE) %>% as.data.frame() 
#make the headers the variable names
 names(temp)  <- heads
```

#kernel density and plots?
##Kernel density and plot for alpha for single FEL result
```{r}
d_alpha_3 <- density(log(temp$alpha), kernel = "gaussian", from = 0) #Kernel density for alpha rate estimation 
#start from 0 and log transform to control outliers

plot(d_alpha_3) #plot the density
lines(d_alpha)
```

## zeiformes atp6 FEL results
```{r}
filepath <- read_json("~/bin/mtDNA_redo/data/FEL/zeiformes-atp8-align-dna.fas.FEL.json") #read in json

heads <- filepath$MLE$headers  %>% unlist() %>% .[c(TRUE,FALSE)] #get headers and ignore header descriptions
```

```{r}
#get MLE contents and make them a data frame
temp <- filepath$MLE$content$`0` %>% unlist %>% matrix(ncol = 6, byrow = TRUE) %>% as.data.frame() 
#make the headers the variable names
 names(temp)  <- heads
```

#kernel density and plots?
##Kernel density and plot for alpha for single FEL result
```{r}
d_alpha_4 <- density(log(temp$alpha), kernel = "gaussian", from = 0) #Kernel density for alpha rate estimation 
#start from 0 and log transform to control outliers

#plot(d_alpha_3) #plot the density
#lines(d_alpha)
```
