---
title: "FEL results preliminary"
author: "sadie"
date: "3/4/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(jsonlite)

library(stringr)

library(dplyr)
library(purrr)
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization
```

# check one alignment
## zeiformes nd1 FEL results
```{r}
filepath <- read_json("~/bin/mtDNA_redo/data/FEL/zeiformes-nd1-align-dna.fas.FEL.json") #read in json

heads <- filepath$MLE$headers  %>% unlist() %>% .[c(TRUE,FALSE)] #get headers and ignore header descriptions
```

```{r}
#get MLE contents and make them a data frame
temp <- filepath$MLE$content$`0` %>% unlist %>% matrix(ncol = 6, byrow = TRUE) %>% as.data.frame() 
#make the headers the variable names
 names(temp)  <- heads
```

#kernel density and plots?
##Kernel density and plot for alpha for single FEL result
```{r}
d_alpha <- density(log(temp$alpha), kernel = "gaussian", from = 0) #Kernel density for alpha rate estimation 
#start from 0 and log transform to control outliers

d_alpha %>% plot() #plot the density 
```
what is the variability of the distribution?
```{r}
d_a_var <- d_alpha$x %>% var()

# t.a <- temp$alpha
# t.a[which(temp$alpha<=0)]  <- 0 
# t.a[which(temp$alpha== -Inf)]  <- 0 
#   log(t.a)%>% var()
```

##Kernel density and plot for beta for single FEL result
```{r}
d_beta <- density(log(temp$beta), kernel = "gaussian", from = 0)  #Kernel density for the beta rate estimation

d_beta %>% plot() #plot the density
```


#check a second alignment:
## acipenseriformes at8 FEL results
```{r}
filepath <- read_json("~/bin/mtDNA_redo/data/FEL/acipenseriformes-atp8-align-dna.fas.FEL.json") #read in json

heads <- filepath$MLE$headers  %>% unlist() %>% .[c(TRUE,FALSE)] #get headers and ignore header descriptions
```

```{r}
#get MLE contents and make them a data frame
temp_2 <- filepath$MLE$content$`0` %>% unlist %>% matrix(ncol = 6, byrow = TRUE) %>% as.data.frame() 
#make the headers the variable names
 names(temp_2)  <- heads
```

#kernel density and plots?
##Kernel density and plot for alpha for single FEL result
```{r}
d_alpha_2 <- density(log(temp_2$alpha), kernel = "gaussian", from = 0)

d_alpha_2 %>% plot()
```
what is the variability of the distribution?
```{r}
d_a_2_var <- d_alpha_2$x %>% var()
```

##Kernel density and plot for beta for single FEL result
```{r}
d_beta_2 <- density(log(temp_2$beta), from = 0)
d_beta_2 %>% plot()
```
#try different ways of calculating the KL or JSD
##using the entropy library:
```{r}
#uses entropy library to calculate Kullback-Leibler divergence (KL) as it is needed to do the JSD
library("entropy")

KL <- KL.plugin(freqs1 = d_alpha$x, freqs2 = d_beta$x)
KL_2 <- KL.plugin(freqs1 = d_alpha_2$x, freqs2 = d_beta_2$x)
#both of these return an error: Vanishing value(s) in argument freqs2!
```
## Using a calculation from stackoverflow:
```{r}
#from stackoverflow https://stackoverflow.com/questions/11226627/jensen-shannon-divergence-in-r
p <- d_beta_2$x
q <- d_alpha_2$x
n <- 0.5 * (p + q)
JS <- 0.5 * (sum(p * log(p / n)) + sum(q * log(q / n))) #returns an NaN
```

##Using a formula from a website:

```{r}
#actually, lets try the way this site does it: https://enterotype.embl.de/enterotypes.html

JSD<- function(x,y) sqrt(0.5 * KLD(x, (x+y)/2) + 0.5 * KLD(y, (x+y)/2))

KLD <- function(x,y) sum(x * log(x/y))

KLD(d_beta_2$x, d_alpha_2$x)
JSD(d_beta_2$x, d_alpha_2$x)

 dist.JSD <- function(inMatrix, pseudocount=0.000001, ...) {
	KLD <- function(x,y) sum(x *log(x/y))
	JSD<- function(x,y) sqrt(0.5 * KLD(x, (x+y)/2) + 0.5 * KLD(y, (x+y)/2))
	matrixColSize <- length(colnames(inMatrix))
	matrixRowSize <- length(rownames(inMatrix))
	colnames <- colnames(inMatrix)
	resultsMatrix <- matrix(0, matrixColSize, matrixColSize)
        
  inMatrix = apply(inMatrix,1:2,function(x) ifelse (x==0,pseudocount,x))

	for(i in 1:matrixColSize) {
		for(j in 1:matrixColSize) { 
			resultsMatrix[i,j]=JSD(as.vector(inMatrix[,i]),
			as.vector(inMatrix[,j]))
		}
	}
	colnames -> colnames(resultsMatrix) -> rownames(resultsMatrix)
	as.dist(resultsMatrix)->resultsMatrix
	attr(resultsMatrix, "method") <- "dist"
	return(resultsMatrix) 
 }


```

###calculate the JSD between alpha and beta
```{r}
d.temp <- dist.JSD(temp %>% select(alpha, beta))

d.temp
```
this last method works and gives me the JSD between the nonsynonymous (beta) and synonymous (alpha) rate distributions. 

so now what if we want the distance for every FEL result?

first, get list of all FEL results:

```{r}
dir <- "~/bin/mtDNA_redo/data/FEL"
FEL_jsons <- list.files(path = dir,
                      pattern = '*FEL.json', recursive = TRUE, full.names = TRUE)
```

 write function for JSD calculation:
```{r}
d.temp <- NULL
JSD.calc <- function(filename){
  
  results <- read_json(filename) #read in json

  heads <- results$MLE$headers  %>% unlist() %>% .[c(TRUE,FALSE)] #get headers and ignore header descriptions
  
  #get MLE contents and make them a data frame
  temp <- results$MLE$content$`0` %>% unlist %>% matrix(ncol = 6, byrow = TRUE) %>% as.data.frame() 
#make the headers the variable names
  names(temp)  <- heads
  
#Kernel density for alpha rate estimation 
#start from 0 and log transform to control outliers
   d_alpha <- density(log(temp$alpha), kernel = "gaussian", from = 0) 
   d_beta <- density(log(temp$beta), kernel = "gaussian", from = 0) 
   
  d_alpha_var<- d_alpha$x %>% var() 
  d_beta_var<- d_beta$x %>% var() 
   
  dist.temp <- dist.JSD(temp %>% select(alpha, beta))
  d.temp <- cbind(d.temp, filename, dist.temp, d_alpha_var, d_beta_var)
  
  return(d.temp)
}
```
check that it works:
```{r}
JSD.calc(paste0(FEL_jsons[1]))
```

this returns a named number of JSD calculations 
```{r}
t <- sapply(FEL_jsons, JSD.calc)
t %>% head()
t <- t %>% t() 
colnames(t) <- c("file", "dist.temp", "d_alpha_var", "d_beta_var")
t.df<- t %>% as.data.frame()
```

```{r}
t.1 <- t.df %>% select(d_alpha_var, d_beta_var)
t.1$d_alpha_var <- t.1$d_alpha_var %>% as.numeric()
t.1$d_beta_var <- t.1$d_beta_var %>% as.numeric()
cleannames<- t.1 %>% rownames()  %>% str_extract_all(., "\\w+(?=-)", simplify = T)
rownames(t.1) <-paste(cleannames[,1], cleannames[,2], sep = "_")

distance <- get_dist(t.1) #computes distance matrix, default is euclidean
fviz_dist(distance, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07")) 
```

cluster with 13 clusters to see is gene groups emerge
```{r}
k2 <- kmeans(t.1, centers = 13, nstart = 25)
fviz_cluster(k2, data = t.1, labelsize = 8)
```
another way of plotting the k means clustering:

```{r}
t.1 %>%
  as_tibble() %>%
  mutate(cluster = k2$cluster,
         gene_order = row.names(t.1)) %>%
  ggplot(aes(d_alpha_var, d_beta_var, color = factor(cluster), label = gene_order)) +
  geom_text()
```




cluster with 25 clusters to see if order clusters emerge
```{r}
k2 <- kmeans(t.1, centers = 25, nstart = 25)
fviz_cluster(k2, data = t.1, labelsize = 8)
```
don't notice anything

### elbow method of finding clusters
from this tutorial still http://uc-r.github.io/kmeans_clustering
```{r}
set.seed(123)

# function to compute total within-cluster sum of square 
wss <- function(k) {
  kmeans(t.1, k, nstart = 10 )$tot.withinss
}

# Compute and plot wss for k = 1 to k = 15
k.values <- 1:15

# extract wss for 2-15 clusters
wss_values <- map_dbl(k.values, wss)

plot(k.values, wss_values,
       type="b", pch = 19, frame = FALSE, 
       xlab="Number of clusters K",
       ylab="Total within-clusters sum of squares")

##or this way:
fviz_nbclust(t.1, kmeans, method = "wss")
```
looks like 4 is the best number of clusters by this method:
```{r}
k2 <- kmeans(t.1, centers = 4, nstart = 25)
fviz_cluster(k2, data = t.1, labelsize = 8)
```
### average silloute method:

```{r}
fviz_nbclust(t.1, kmeans, method = "silhouette")
```

looks like 5 is the best number of clusters by this method:
```{r}
k2 <- kmeans(t.1, centers = 5, nstart = 25)
fviz_cluster(k2, data = t.1 labelsize = 8)
```
### gap methond
```{r}
set.seed(123)
gap_stat <- clusGap(t.1, FUN = kmeans, nstart = 25,
                    K.max = 10, B = 50)
# Print the result
print(gap_stat, method = "firstmax")
```

```{r}
fviz_gap_stat(gap_stat)
```

## heirarchical clustering
```{r}

```


take that clean it up so you have gene, order, alpha/beta JSD.

```{r}
# cleanup <- str_extract_all(names(t), "\\w+(?=-)", simplify = T)
# f <- tibble(order = cleanup[,1], gene = cleanup[,2], JSD_alpha_beta = t)
```

stopped here 
```{r}
# get_dist()
# fviz_dist()
```

